---
title: "Rt"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Stochastic model - Gillespie's algorithm
### SEPAIR model 

#### Install packages
```{r}
# devtools::load_all()
# devtools::install()
# devtools::install_github("kimfinale/pfilterCOVID")
```

### Initial condition
```{r, eval=FALSE}
library(pfilterCOVID)
# beta is updated by R0
THETA <- list(beta = 0.3, R0 = 1.2, epsilon = 1/2.5, delta = 1/5, 
            gamma = 1/2.5, bp = 1, ba = 1, fa = 0.3, fd = 1, I0 = 1e2, 
            time_dep_Rt = FALSE)
# this changes the default y0 in the pfilterCOVID package
THETA$Y0 <- c(S = 1e7 - THETA$I0, E = 0, P = 0, A = 0, 
              I = THETA$I0, R = 0, CE = 0, CI = 10, CR = 0)
THETA$tend <- 200
usethis::use_data(THETA)
```

### Run the ODE version
```{r}
# tstamp <- format(Sys.time(), "%Y%m%dT%H%M%S")
library(pfilterCOVID)
library(ggplot2)
theme_set(theme_bw())
params <- THETA
tend <- params$tend

params$I0 <- 100
params$Y0 <- c(S = 1e7 - params$I0, E = 0, P = 0, A = 0, 
              I = params$I0, R = 0, CE = 0, CI = 0, CR = 0)

tstamp <- format(Sys.time(), "%Y%m%d")
times <- seq(0, tend, by = 1) # daily output for 150 days
library(deSolve)
library(tidyverse)

params$time_dep_Rt <- TRUE
ode(y = params$Y0, times = times, func = ode_sepair, parms = params) %>% 
  as.data.frame() -> out 

out$daily_infected <- c(0, diff(out$CE)) 
out$daily_symptom <- c(0, diff(out$CI)) 
out$daily_confirmed <- c(0, diff(out$CR))
out$daily_Rt <- sapply(0:tend, function(x) get_Rt(x))

# saveRDS(out, paste0("outputs/ode_I0_100_", tstamp, ".rds"))

out_long <- out %>% pivot_longer(-time) 
out_long$name <- factor(out_long$name,
                        levels = c("S", "E", "P", "A", "I", "R", 
                                  "CE", "CI", "CR", "daily_infected", 
                                  "daily_symptom", "daily_confirmed", "daily_Rt"))
out_long %>%
  filter(name == "daily_infected" | 
           name == "daily_symptom" | name == "daily_confirmed") %>%
  ggplot(aes(x = time, y = value, color = name)) +
  geom_line(size = 1.2) +
  labs(x='Time (day)', y = 'Number of individuals', color = "") + 
  theme_grey(base_size = 16) 
  # facet_wrap(vars(name), nrow = 2, scales = "free_y")
```

### Run the Gillespie's direct method 
This is to create a stochastic data set based on the Gillespie's direct method 
```{r}
# library(pfilterCOVID)
# library(tidyverse)
set.seed(42)
tstart <- Sys.time()
# tstamp <- format(Sys.time(), "%Y%m%d")
nrun <- 20

res <- gillespie_run(func = gillespie_sepair, 
                     tend = tend, 
                     nrun = nrun, 
                     y0 = params$Y0, 
                     params = params, 
                     report_dt = 1)
Sys.time() - tstart

daily_infected <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))
daily_symptom <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))
daily_confirmed <- data.frame(matrix(0, nrow = tend + 1, ncol = nrun))

for (i in seq_len(length(res))) {
  inf <- diff(res[[i]]$CE)
  onset <- diff(res[[i]]$CI)
  conf <- diff(res[[i]]$CR)
  daily_infected[1:(length(inf)+1), i] <- c(0, inf)
  daily_symptom[1:(length(onset)+1), i] <- c(0, onset)
  daily_confirmed[1:(length(conf)+1), i] <- c(0, conf)
}

df <- daily_infected
df$time <- 0:tend
df %>% pivot_longer(cols = -time) -> df
dfrm <- data.frame(time = 0:tend, dinf = rowMeans(daily_infected))

plt <- ggplot() +
  geom_line(data = df, aes(time, value, group = name), 
            color = "grey80") +
  geom_line(data = out, aes(time, daily_infected), # ODE model output
            color = "darkred",
            size = 1, inherit.aes = FALSE) + 
  geom_line(data = dfrm, aes(time, dinf), # ODE model output
            linetype = "dotted",
            size = 1, inherit.aes = FALSE) + 
  labs(x="Time", y="Daily infected cases", color="") + 
  geom_segment(aes(x=c(0,0,0), xend=c(5,5,5), y=c(2000,1800,1600),
                yend=c(2000,1800,1600)), size=c(1,1,1), 
               color = c("grey80", "darkred", "black"),
               linetype = c("solid","solid","dotted")) +
  geom_text(aes(x=c(7,7,7), y=c(2000,1800,1600)), size=c(4,4,4), 
               hjust = c(0,0,0), vjust = c(0.5,0.5,0.5), 
               label = c("Stochastic simulations", "ODE", "Mean of the stochastic simulations"))
# plt            
# ggsave("plots/ODE_stoch_IO_100_infected.png", plt)

stoch <- list()
stoch$daily_confirmed <- daily_confirmed
stoch$daily_symptom <- daily_symptom
stoch$daily_infected <- daily_infected
stoch$daily_Rt <- sapply(0:tend, function(x) get_Rt(x))

# saveRDS(stoch, paste0("outputs/stoch_I0_100_", tstamp, ".rds"))
```

### Apply a detection rate
This is to create a data set in which the detection rate is not 100%.
```{r}
rr <- nrow(stoch$daily_confirmed)
cc <- ncol(stoch$daily_confirmed)
detected <- data.frame(matrix(0, nrow = rr, ncol = cc))
stoch_partial_obs <- list()
probs <- seq(0.2, 0.8, 0.1)
stoch_partial_obs$probs <- probs
set.seed(42)
for (k in 1:length(probs)) {
  prob <- probs[k]
  for (i in 1:rr) {
     for (j in 1:cc){
       size <- stoch$daily_confirmed[i, j]
       # message(paste0("i = ", i, ", j = ", j, ", case = ", size, "\n"))
       if (!is.na(size) & size > 0){
         detected[i, j] <- rbinom(1, size = size, prob = prob)
       }
    }
  }
  stoch_partial_obs$daily_detected[[k]] <- detected
}

# saveRDS(stoch_partial_obs, paste0("outputs/stoch_I0_100_partial_obs_", tstamp, ".rds"))
```

## Total dataset
The data set that contains all the necessary information: Rt trajectory, ODE
and stochastic model result
```{r}
Rt_dataset <- list()
Rt_dataset$README <- "To see if potential bias of Rt estimation based on the confirmation data set, especially in the most recent time, is reduced or eliminated when the number of cases are large"
Rt_dataset$ode <- out
Rt_dataset$stoch_perfect_obs <- stoch
Rt_dataset$stoch_partial_obs <- stoch_partial_obs
Rt_dataset$Rt <- sapply(0:tend, function(x) get_Rt(x))
Rt_dataset$time <- out$time
Rt_dataset$params <- params
Rt_dataset$y0 <- params$Y0
# saveRDS(Rt_dataset, paste0("outputs/Rt_dataset_I0=100_", tstamp, ".rds"))
```

### Plot: Rt, infection/confirmation/stoch outputs
```{r}
# Rt data set; Rt_dataset_I0=100_20220105.rds, Rt_dataset_I0=10_20220105.rds
tstamp <- "20220105"
dat <- readRDS(paste0("outputs/Rt_dataset_I0=100_", tstamp, ".rds"))

library(ggplot2)
plt <- ggplot() +
  geom_line(aes(x=dat$time, y = dat$Rt)) +
  labs(x="time", y=expression(R[t]))

# ggsave("plots/Rt.png", plt)

library(tidyverse)
oderes <- dat$ode[, c("time", "daily_infected", "daily_confirmed")]
oderes %>% pivot_longer(cols = -time) -> oderes

stochres <- dat$stoch_perfect_obs$daily_confirmed[, 1:20]
stochres$time <- dat$time

stochres %>% pivot_longer(cols = -time) -> stochres

plt <- ggplot() +
  # geom_line(data = stochres, aes(time, value, group=name), color="grey50") +
  geom_line(data = oderes, aes(time, value, color=name), size=1.2) + 
  labs(x="time", y="", color="") + 
  theme(legend.position="top")
   
# ggsave("plots/daily_inf_conf.png", plt)
#-------------------------------------------------------------------
# data set generation for Rt presentation 
 
df <- data.frame(Day = dat$time, 
                 Rt = round(dat$ode$daily_Rt, 1), 
                 Infected = round(dat$ode$daily_infected), 
                 Confirmed = round(dat$ode$daily_confirmed))
data.table::fwrite(df, "outputs/data.csv")

# Serial interval for the model from Mathematica
mean <- 6.25 # mean_si
std <- 4.14578 # std_si
get_gamma_params <- function(mean, std) {
  # k*theta*theta = variance
  # k*theta = mean
  theta <- std*std/mean # scale
  k <- mean/theta # shape
  return(list(k = k , theta = theta))
}
gampar <- get_gamma_params(mean = mean, std = std)
plot(0:20, dgamma(0:20, shape = gampar$k, scale = gampar$theta), xlab = "Day", ylab = "Prob")
d <- data.frame(Day = 0:20, Prob = dgamma(0:20, shape = gampar$k, scale = gampar$theta))
plt <- ggplot(d, aes(Day, Prob)) +
  geom_point() + 
  labs(title = expression(Generation~interval~omega))

ggsave(paste0("plots/si.png"), plt, width=3.4*1.6, height=2.7*1.6, units="in")

calc_Rt

df$Infected[101] / sum(df$Infected[91:100] * dgamma(10:1, shape = gampar$k, scale = gampar$theta))  


df$Infected[101] / sum(df$Infected[81:100] * dgamma(20:1, shape = gampar$k, scale = gampar$theta))  


df2 <- data.frame(Day = 0:20, wt = round(dgamma(0:20, shape = gampar$k, scale = gampar$theta),3))
data.table::fwrite(df2, "outputs/wt.csv")
#-----------------------------------------------------------------
```

### Rt estimation using the particle filter 
Particle filtering - particle_filter
Expect errors since backward sampling is not in place
```{r}
# library(pfilterCOVID)
set.seed(1)
tstamp <- "20220105"
I0 <- 100
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
# determine the data type to model (infection vs. confirmation)
dtype <- "confirmation"
d <- dat$ode[, c("time", "daily_infected")]
if (dtype == "confirmation") {
  d <- dat$ode[, c("time", "daily_confirmed")]
}

dat$params$time_dep_Rt <- FALSE # this must be set FALSE to produce an estimate

params <- dat$params
#standard deviation of the prior distribution
params[["betavol"]] <- 0.2
tbegin <- Sys.time()

pf <- pfilter(
  params = params,
  y = params$Y0,
  data = d,
  data_type = dtype,
  npart = 1e4,
  tend = nrow(d),
  dt = 0.1,
  error_pdf = "pois",
  systematic_resampling = FALSE,
  backward_sampling = TRUE,
  stoch = FALSE)

Sys.time() - tbegin

# Rt <- readRDS("outputs/Rt_default.rds")
epsilon <- params[["epsilon"]]
delta <- params[["delta"]]
gamma <- params[["gamma"]]
bp <- params[["bp"]] # relative infectiousness of pre-symptomatic state
ba <- params[["ba"]] # relative infectiousness of asymptomatic state
fa <- params[["fa"]] # fraction of asymptomatic state

durP <- (1 / delta - 1 / epsilon)
durI <- (1 / gamma)
R0_dur <- ((1 - fa) + fa * ba) * durI + bp * durP

med <- apply(pf$beta_filtered, 1, quantile, probs = c(0.5))
# plot(dat$time + 2, dat$daily_Rt, type="l", ylim=c(0, max(upper * R0_dur)))
plot(dat$ode$time + 2, dat$ode$daily_Rt, type="l")
lines(med * R0_dur, lwd=2, col=3)
```

### Fitting the infection or confirmation time series from the ODE model
This uses the extract_trace function that calls pfilter function and also executes backward sampling
#### PF 
```{r}
library(pfilterCOVID)
library(ggplot2)
theme_set(theme_bw())

tstamp <- "20220105"
dat <- readRDS(paste0("outputs/Rt_dataset_I0=100_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE
params[["betavol"]] <- 0.2 

# increase the number to improve accuracy
nrep <- 1e2
npart <- 1e2
dt <- 0.2
library(parallel)
library(doParallel)

ncores <- detectCores() 
dtype <- "confirmation"
# ODE with perfect observation
d <- data.frame(date = dat$time,
               daily_infected = round(dat$ode$daily_infected))
if (dtype == "confirmation") {
  d <- data.frame(date = dat$time,
                  daily_confirmed = round(dat$ode$daily_confirmed))
}

set.seed(42)
cl <- makeCluster(getOption("cl.cores", 2))
doParallel::registerDoParallel(cl)

pf <- foreach(i = 1:nrep, .packages = c("pfilterCOVID"), .inorder = F) %dopar% {
 extract_trace(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = dtype,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
}
parallel::stopCluster(cl)
  
parset_chr <- paste0(dtype, "_I0=", params$I0, "_npart=", npart,
           "_nrep=", nrep, "_dt=", dt, "_", tstamp)
# saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
  
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
  
df <- cbind(Rt_quantile, d)
col_fill <- "#1F618D"
col_dat <- "grey70"
df$Rt <- dat$ode$daily_Rt

plt <- ggplot(df, aes(x = date)) +
    geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
    geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
    geom_line(aes(y = `50%`), color = col_fill, size = 1.2) +
    geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
    geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
    labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
plt  
# ggsave(paste0("plots/", parset_chr, ".png"), plt)
```
#### EpiEstim
```{r}

```

#### Figure 
```{r}

```


### Fitting confirmation time series from the stochastic model 
```{r}
# library(pfilterCOVID)
library(ggplot2)
theme_set(theme_bw())
tstamp <- "20220105"
I0 <- 100 
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e2
npart <- 1e3
dt <- 0.2
betavol <- 0.2

library(parallel)
library(doParallel)
ncores <- detectCores() 

dtype <- "confirmation"
for (k in 1:20) {
  d <- data.frame(date = dat$time,
                 daily_confirmed = round(dat$stoch_perfect_obs$daily_confirmed[,k]))
  
  tailsum <- sum(tail(d$daily_confirmed))
  message(paste0("k = ", k, ", tailsum = ", tailsum))
  
  if (tailsum > 0) {
    tbegin <- Sys.time()
    
    set.seed(42)
    cl <- makeCluster(getOption("cl.cores", 2))
    doParallel::registerDoParallel(cl)
    
    params[["betavol"]] <- betavol
    
    pf <- foreach(i = 1:nrep, .packages = c("pfilterCOVID"), .inorder = F) %dopar% {
      extract_trace(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = dtype,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
    }
    parallel::stopCluster(cl)
      
    telapsed <- Sys.time() - tbegin
    message(paste0("time elapsed = ", telapsed))
    
    parset_chr <- 
        paste0(dtype, "_I0=", params$I0, "_betavol=", params[["betavol"]], "_npart=", npart, "_nrep=", nrep, "_dt=", dt, "_", "k=", k, "_", tstamp)
      
    saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
      
    pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
    Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
    Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
      
    df <- cbind(Rt_quantile, d)
    col_fill <- "#1F618D"
    col_dat <- "grey70"
    df$Rt <- dat$ode$daily_Rt
    
    plt <- 
      ggplot(df, aes(x = date)) +
      geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
      geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
      geom_line(aes(y = `50%`), color = col_fill, size = 1) +
      geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
      geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
      labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
        
    ggsave(paste0("plots/", parset_chr, ".png"), plt)
  }
}
```

#### EpiEstim
```{r}
#deconvolution
```

#### RMSE comparison
```{r}
```

#### Figure
```{r}

```

### Fitting confirmation time series from the stochastic model
This is designed to explore how the rolling average may help to identify the overall pattern by smoothing short-term fluctuations
```{r}
# library(pfilterCOVID)
library(ggplot2)
theme_set(theme_bw())

tstamp <- "20220105"
I0 <- 10
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e2
npart <- 1e3
dt <- 0.2
betavol <- 0.2

library(parallel)
library(doParallel)

ncores <- detectCores() 
dtype <- "confirmation"

roll_window <- 7
for (k in 1:20) {
  dd <- dat$stoch_perfect_obs$daily_confirmed[,k]
  roll_case <- 
    data.table::frollmean(dd, n=roll_window, align="center")
  # replace the missing values, from rolling mean, with the original values just for convenience. This should not affect the estimates as they are the at the very beginning or the very end
  roll_case[1:3] <- dd[1:3] # include original data
  roll_case[(length(dd)-3):length(dd)] <- dd[(length(dd)-3):length(dd)] # include original data
  
  d <- data.frame(date = dat$time,
                 daily_confirmed = round(roll_case))
  
  tailsum <- sum(tail(d$daily_confirmed))
  message(paste0("k = ", k, ", tailsum = ", tailsum))
  
  if (tailsum > 0) {
    tbegin <- Sys.time()
    
    set.seed(42)
    cl <- makeCluster(getOption("cl.cores", 2))
    doParallel::registerDoParallel(cl)
    
    params[["betavol"]] <- betavol
    
    pf <- foreach(i = 1:nrep, .packages = c("pfilterCOVID"), .inorder = F) %dopar% {
      extract_trace(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = dtype,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
    }
    parallel::stopCluster(cl)
      
    telapsed <- Sys.time() - tbegin
    message(paste0("time elapsed = ", telapsed))
    
    parset_chr <- 
        paste0(dtype, "_I0=", params$I0, "_betavol=", params[["betavol"]], "_npart=", npart, "_nrep=", nrep, "_dt=", dt, "_", "k=", k, "_rollwd=", roll_window, "_", tstamp)
      
    saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
      
    pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
    Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
    Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
      
    df <- cbind(Rt_quantile, d)
    col_fill <- "#1F618D"
    col_dat <- "grey70"
    df$Rt <- dat$ode$daily_Rt
    
    plt <- ggplot(df, aes(x = date)) +
        geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
        geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
        geom_line(aes(y = `50%`), color = col_fill, size = 1) +
        geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
        geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
        labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
        # scale_y_continuous(limits=c(0,3))
    
    # plt  
    ggsave(paste0("plots/", parset_chr, ".png"), plt)
  }
}  
```

### Fitting confirmation time series from the stochastic model with imperfect
observation
```{r}
# library(pfilterCOVID)
library(ggplot2)
theme_set(theme_bw())
tstamp <- "20220105"
I0 <- 100 
dat <- readRDS(paste0("outputs/Rt_dataset_I0=", I0, "_", tstamp, ".rds"))
params <- dat$params
params$time_dep_Rt <- FALSE

nrep <- 1e2
npart <- 1e3
dt <- 0.2
betavol <- 0.2

library(parallel)
library(doParallel)
ncores <- detectCores() 

dtype <- "confirmation"
detect_prob <- 0.3
# id <- which(dat$stoch_partial_obs$probs == 0.3) # bug when 0.3 is applied...
id <- grep(detect_prob, dat$stoch_partial_obs$probs)
# 1 through 7 indicating detection prob ranging from 0.2 to 0.8
for (k in 1:20) {
  d <- data.frame(date = dat$time,
                  daily_confirmed = round(dat$stoch_partial_obs$daily_detected[[id]][,k]))
  
  tailsum <- sum(tail(d$daily_confirmed))
  message(paste0("k = ", k, ", tailsum = ", tailsum))
  
  if (tailsum > 0) {
    tbegin <- Sys.time()
    
    set.seed(42)
    cl <- makeCluster(getOption("cl.cores", 2))
    doParallel::registerDoParallel(cl)
    
    params[["betavol"]] <- betavol
    
    pf <- foreach(i = 1:nrep, .packages = c("pfilterCOVID"), .inorder = F) %dopar% {
      extract_trace(params = params,
                    y = params$Y0,
                    data = d,
                    data_type = dtype,
                    npart = npart,
                    tend = nrow(d),
                    dt = dt,
                    error_pdf = "pois",
                    negbin_size = 15,
                    backward_sampling = TRUE,
                    stoch = FALSE)
    }
    parallel::stopCluster(cl)
      
    telapsed <- Sys.time() - tbegin
    message(paste0("time elapsed = ", telapsed))
    
    parset_chr <- 
        paste0(substr(dtype,1,4), "_I0=", params$I0, "_betavol=", params[["betavol"]], "_np=", npart, "_nrep=", nrep, "_dt=", dt,
               "_", "prob=", detect_prob, "_", "k=", k, "_", tstamp)
      
    saveRDS(pf, paste0("outputs/pf_", parset_chr, ".rds"))
      
    pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
    Rt_est <- as.data.frame(sapply(pf, function(x) x[, "Rt"]))
    Rt_quantile <- as.data.frame(t(apply(Rt_est, 1, function(x) quantile(x, pr))))
      
    df <- cbind(Rt_quantile, d)
    col_fill <- "#1F618D"
    col_dat <- "grey70"
    df$Rt <- dat$ode$daily_Rt
    
    plt <- 
      ggplot(df, aes(x = date)) +
      geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = col_fill, alpha = 0.5) +
      geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = col_fill, alpha = 0.8) +
      geom_line(aes(y = `50%`), color = col_fill, size = 1) +
      geom_line(aes(x = date + 1, y = Rt), size = 1, linetype = "dashed") +
      geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted")+
      labs(title = parset_chr, y = expression(italic(R)[italic(t)]), x = "")
        
    ggsave(paste0("plots/", parset_chr, ".png"), plt)
  }
}
```


### Fitting confirmation time series from the stochastic model
While generation time (aka, generation interval) is used to correctly reproduce
the epidemic curve, serial interval, time interval between symptom onsets of 
successive cases, is often used instead because of its easier estimation.
For this, we can set the serial interval differently for the EpiEstim estimation

EpiEStim with shortened serial interval.

```{r}

```



